Massively Parallel Monte Carlo Methods for Discrete Linear and Nonlinear Systems

### 2012-10-24 doc/prelim/presentation/prelim_presentation.tex
presentation updates

\date{\today}
\title{Massively Parallel Monte Carlo Methods for Linear and Nonlinear
Problems}
\begin{document}
\maketitle

%%---------------------------------------------------------------------------%%
\begin{frame}{Introduction}
  \begin{itemize}
    \item Predictive modeling and simulation capability yields nuclear
      system designs that are safer and better performing.
    \item Modern work focused on this task leverages multiple physics
      simulation (CASL, NEAMS).
    \item New hardware drives algorithm development (petascale and
      exascale).
    \item Monte Carlo methods have the potential to provide great
      improvements that permit finer simulations and better mapping to
      future hardware.
    \item A set of massively parallel Monte Carlo methods is proposed
      to advance multiple physics simulation on contemporary and
      future leadership class machines.
  \end{itemize}
\end{frame}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Hardware-Based Motivation}
  \begin{itemize}
    \item Modern hardware is moving in two directions: lightweight
      machines and heterogeneous machines characterized by low power
      and high concurrency.
    \item High concurrency and low cost units means a higher potential
      for both soft and hard failures.
    \item Monte Carlo methods bury soft failures within the variance
      of the tallies while hard failures are high variance events.
    \item New machines will also be memory restricted with a continued
      decrease of memory/FLOPS predicted.
    \item Compared to conventional methods, we will show that Monte
      Carlo methods offer a memory savings.
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Research Outline}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Conventional Solution Methods for Linear Problems}

%%---------------------------------------------------------------------------%%
\begin{frame}{Linear Problem Preliminaries}

%%---------------------------------------------------------------------------%%
\begin{frame}{Stationary Methods}

@@

>>>>

\date{\today} 
\title{Massively Parallel Monte Carlo Methods for Discrete Linear and
  Nonlinear Systems} 
\begin{document}
\maketitle

%%---------------------------------------------------------------------------%%
\begin{frame}{Introduction}
  \begin{itemize}
    \item Predictive modeling and simulation ehances engineering
      capability to yield nuclear system designs that are safer and
      better performing.
    \item Modern work focused on this task leverages multiple physics
      simulation (CASL, NEAMS).
    \item New hardware drives algorithm development (petascale and
      exascale).
    \item Monte Carlo methods have the potential to provide great
      improvements that permit finer simulations and better mapping to
      future hardware.
    \item A set of massively parallel Monte Carlo methods is proposed
      to advance multiple physics simulation on contemporary and
      future leadership class machines.
  \end{itemize}
\end{frame}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Hardware-Based Motivation}
  \begin{itemize}
    \item Modern hardware is moving in two directions: lightweight
      machines and heterogeneous machines characterized by low power
      and high concurrency.
    \item High concurrency and low cost units means a higher potential
      for both soft and hard failures.
    \item Monte Carlo methods bury soft failures within the variance
      of the tallies while hard failures are high variance events.
    \item New machines will also be memory restricted with a continued
      decrease of memory/FLOPS predicted.
    \item Compared to conventional methods, we aim show that Monte
      Carlo methods offer a memory savings.
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Research Outline}
  \begin{itemize}
    \item Parallelization of classic Monte Carlo methods.
    \item Parallel strategies taken from modern reactor physics
      methods in Monte Carlo.
    \item Research is required to explore how domain decomposition
      patterns and the discrete system properties are related.
    \item Research is required to explore how these methods perform
      in modern multiple physics simulations where strong
      nonlinearities are present.
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{The Linear Problem}

  We will seek solutions of the general linear problem:

  \[
  \ve{A} \ve{x} = \ve{b}\:,
  \]
  \[
  \ve{A} \in \mathbb{R}^{N \times N},\ \ve{A} : \mathbb{R}^{N}
  \rightarrow \mathbb{R}^{N},\ \ve{x} \in \mathbb{R}^N,\ \ve{b} \in
  \mathbb{R}^N\:.
  \]

  We will assert that $\ve{A}$ is \textit{nonsingular}. The solution
  is then:
  \[
  \ve{x} = \ve{A}^{-1}\ve{b}\:.
  \]

  We can also define the residual of the system as:
  \[
  \ve{r} = \ve{b} - \ve{A}\ve{x}\:,
  \]
  where $\ve{r}=\ve{0}$ when an exact solution is found.

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Stationary Methods}
  General stationary methods are formed by splitting the linear
  operator:
  \[
  \ve{A} = \ve{M} - \ve{N}\:.
  \]

  We can then reformulate the linear system as:
  \[
  \ve{x} = \ve{M}^{-1}\ve{N}\ve{x} + \ve{M}^{-1}\ve{b}\:.
  \]

  We identify $\ve{M}^{-1}\ve{N}$ as the \textit{iteration matrix},
  $\ve{H}$, and apply iteration indices to the solution vectors to
  give an iterative method.
  \[
  \ve{x}^{k+1} = \ve{H}\ve{x}^{k} + \ve{c}\:.
  \]

  The qualities of the iteration matrix dictate convergence. Defining
  $\ve{e}^k = \ve{x}^k-\ve{x}$ as the error at the $k^{th}$ iterate
  gives a recurrence relation:
  \[
  \ve{e}^{k+1} = \ve{H} \ve{e}^k\:,
  \]

  By diagonalizing $\ve{H}$ to extract its Eigenvalues, we find that 
  \[
  ||\ve{e}^{k}||_2 = \rho(\ve{H})^k ||\ve{e}^0||_2\:,
  \]
  and therefore we bound $\ve{H}$ by $\rho(\ve{H}) < 1$ for convergence.

@@

  We choose a search subspace $\mathcal{K}$ and a search subspace
  $\mathcal{L}$ and determine the solution to the linear problem by
  extracting the solution from the search subspace:
  \[
  \tilde{\ve{x}} = \ve{x}_0 +
  \boldsymbol{\delta},\ \boldsymbol{\delta} \in \mathcal{K}\:,
  \]
  and by constraining it with the constraint subspace:
  \[
  \langle \tilde{\ve{r}},\ve{w} \rangle = 0,\ \forall \ve{w} \in
  \mathcal{L}\:.
  \]
